A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-15 tex/codatadp.tex
more

\newcommand\landg{L\,\&\,G}
\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}

\begin{table}[tbh!]
\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
\\[1pt] %%% TYPESETTING HACK
\midrule
%$\enatT$
\\[-9pt] %%% TYPESETTING HACK
No decision procedure
  & 800 & 922 & 450 & 2172 \\
Codatatypes without bisimilarity
  & 810 & 936 & 470 & 2216 \\
Codatatypes with bisimilarity
  & 811 & 939 & 472 & 2222 \\
Datatypes without cyclicity
  & 815 & 947 & 469 & 2231 \\
Datatypes with cyclicity
  & 818 & 953 & 470 & 2241 \\
Full decision procedure
  & 840 & 970 & 490 & 2300 \\[\jot]
Total number of goals
  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
\end{tabular}\end{center}
\caption{Number of solved goals for the three benchmark suites}
\label{tab:bench}
\end{table}

To evaluate the decision procedure, we generated problems from existing
Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
included the theory files from the Isabelle distribution (the Distro) and the \emph{Archive
of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
within the supported fragment. We also included two unpublished theories by
Andreas Lochbihler and Peter Gammie (\landg). We complemented the benchmarks
with some theories about various list and tree data structures, to also test
the datatypes.
The theory files were selected before running the experiments.
Our experimental data is publicly available \cite{our-eval-data}.

@@

\def\ackname{Acknowledgment}
\paragraph{\ackname.}
We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
blah blah. \textbf{TODO FIXME}
%
Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
codatatype or at least endured
%tolerated?
it, both of which we are thankful for.
%
Andreas Lochbihler and Peter Gammie shared their private
theories with us so we could include them in the benchmarks.
Andrei Popescu helped clarify our thoughts regarding the axiomatization of
codatatypes. Dmitriy Traytel took part in discussions about special
cases.
%
Blanchette's research was partially supported by the Deutsche
Forschungs\-gemein\-schaft (DFG) project
\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).

>>>>

\newcommand\gandl{G\&L}
\newcommand\HD[1]{\hbox to2.5em{\hfill#1\hfill}}

\begin{table}[tbh!]
\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
  & \HD{Distro} & \HD{AFP} & \HD{\gandl} & All
\\[1pt] %%% TYPESETTING HACK
\midrule
%$\enatT$
\\[-9pt] %%% TYPESETTING HACK
No decision procedure
  & 800 & 922 & 450 & 2172 \\
Codatatypes without bisimilarity
  & 810 & 936 & 470 & 2216 \\
Codatatypes with bisimilarity
  & 811 & 939 & 472 & 2222 \\
Datatypes without cyclicity
  & 815 & 947 & 469 & 2231 \\
Datatypes with cyclicity
  & 818 & 953 & 470 & 2241 \\
Full decision procedure
  & 840 & 970 & 490 & 2300 \\[\jot]
Total number of goals
  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
\end{tabular}\end{center}
\caption{Number of solved goals for the three benchmark suites}
\label{tab:bench}
\end{table}

To evaluate the decision procedure, we generated problems from existing
Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
included the theory files from the Isabelle distribution (the Distro) and the \emph{Archive
of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
within the supported fragment. We also included two unpublished theories by
Peter Gammie and Andreas Lochbihler (\gandl). We complemented the benchmarks
with some theories about various list and tree data structures, to also test
the datatypes.
The theory files were selected before running the experiments.
Our experimental data is publicly available \cite{our-eval-data}.

@@

\def\ackname{Acknowledgment}
\paragraph{\ackname.}
We owe a great debt to Clark Barrett and Cesare Tinelli for datatype case blah
blah blah. \textbf{TODO FIXME}
%
Our present and former bosses, Viktor Kuncak, Stephan Merz, Tobias Nipkow,
Cesare Tinelli, and Christoph Weidenbach, have either encouraged the research on
codatatype or at least endured
%tolerated?
it, both of which we are thankful for.
%
Peter Gammie and Andreas Lochbihler shared their private
theories with us so we could include them in the benchmarks.
Andrei Popescu helped clarify our thoughts regarding the axiomatization of
codatatypes. Dmitriy Traytel took part in discussions about special
cases.
%
Blanchette's research was partially supported by the Deutsche
Forschungs\-gemein\-schaft (DFG) project
\relax{Hardening the Hammer} (grant Ni\,491\slash 14-1).
